What's in a Domain? Anaylsis of URL Features

### 2022-10-02 paper/paper.tex
Paper update

The Uniform Resource Locator (URL) is now a ubiquitous element of modern life. 
We use them to advertise and locate businesses, to retrieve news and media and 
to interact with other people across a broad of social applications. It should
be no surprise that processing URLs is key component of many tasks that involve
analysing internet data. In many applications the URL is the central piece of 
information available because the demands of the task require immediate analysis.
Prominent examples include security analysis of potenital phishing attacks\cite{Basnet2014}, 
detection of potential spam emails \cite{}, and applications to online advertsing
including anticipation of conversions \cite{Qiu2020} or contextual analysis of
the content\cite{Kan2005,Baykan2009} or language\cite{Baykan2013} of webpages.

@@

Multiple data science and machine learning applications on internet data 
Write a literature review \cite{Vazhayil2018}
\cite{Canali2011}
\cite{Ayes2019}
\cite{Mamun2016}
\cite{Li2020}
\cite{Basnet2012}
\cite{Basnet2014}
\cite{Tupsamudre2019}
URLnet learns word and charcater embedings \cite{Le2018}
\cite{Baykan2009}
\cite{Kan2005}
\cite{Hernandez2012}
\cite{Chung2010}
\cite{Qiu2020}
\cite{Vaishnavi2021}
\cite{Xu2021}
\cite{Rhea2022}
\cite{Ma2009}
\cite{Verma2017}
\cite{Sountharrajan2020}

\cite{Murthy2022}
Classify Web pages based on URLs from the 4 Universities WebKB dataset. Comparing features derived from segmentation
of the URL to features derived from either link anchor text or the web page document itself \cite{Kan2004}.

Common approaches to feature engineering on URLs include string patterns and regular expressions to
identify key sequences \cite{Kan2004}, n-gram or bag of words models \cite{Baykan2009}, 
the generation of task specific embedding vectors\cite{Le2018,Qiu2020} and the
usage of domain name servers or registrars for ancillary information \cite{}. Some authors have 
empahsized that URLs are sequences of characters and require features that respect this ordering\cite{Le2018}.
However, the ordering of key features within URLs is predominantly fixed.

@@

The features generated by the url2feature package are summaize in Table \ref{tab:features}

\begin{table*}
\caption{URL Features}
\label{tab:features}
\begin{tabular}{|l|l|l|l|}
\toprule
Group         &Feature              &Type        &Definition  \\
\midrule
Protocol      &protocol\_name        &Category    &The exact protocol name    \\
              &protocol\_type        &Category    &Protocol categorised by purpose    \\
              &protocol\_exists      &Boolean     &Flag indicating presence of protocol   \\
\midrule
Host          &host\_is\_ip           &Boolean     &Flag indicating if the host is an IP address    \\
              &ext\_type             &Category    &TLD extension categorised by purpose    \\
              &ext\_freq             &Boolean     &TLD extension frequency in internet traffic   \\
\midrule
Path          &path\_depth           &Numeric     &The number of forward slashes    \\
              &path\_words           &Numeric     &Count of distinct words    \\
\midrule
File          &file\_ext             &Category    &The exact file extension e.g. "exe" or "php"    \\
              &file\_type            &Category    &File extension categorised by function \& purpose    \\
              &file\_ext\_exists      &Boolean     &Flag indicating presence of protocol   \\
\midrule
Params        &name                 &Category    &The exact protocol name    \\
              &type                 &Category    &Protocol categorised by purpose    \\
              &exists               &Boolean     &Flag indicating presence of protocol   \\
\bottomrule
\end{tabular}
\end{table*}

@@

The data used in this study was collected from a collction of publications that 
present internet classification problems containing URLs as a key, or sole feature.
The IS Malcious URL classification dataset contains both malware and phishing URLS 
that need to be discriminated against benign URLS\cite{Mamun2016}.
The world wide web knowledge base (WebKb) 4 Universities data set contains a wide
range of university URLs categorised into multiple topic categories\cite{Craven1998}.
These 3 URL classification problems are summarised in Table \ref{tab:data}

\begin{table}
\caption{Datasets}
\label{tab:data}
\begin{tabular}{|l|r|r|}
\toprule
Dataset        &Records        &Classes  \\
\midrule
Malware        &               &2        \\
Phishing       &               &2        \\
WebKb 4Uni     &               &7        \\
\bottomrule
\end{tabular}
\end{table}

>>>>

The Uniform Resource Locator (URL) is a ubiquitous element the digital world. 
We use them to advertise and locate businesses, to retrieve news and media and 
to interact with other people across a broad of social applications.
Processing URLs is key component of many tasks that involve
analysing internet data. In many applications the URL is the central piece of 
information available because the demands of the task require immediate analysis.
In applications like malicious wesite detection the URL needs to be processed with
in a rapid and efficient manner to provide utility\cite{Ma2009b}.
Common use cases for URL centered analysis include security analysis of potenital phishing 
attacks\cite{Basnet2014,Mamun2016,Vazhayil2018}, and identification of pages 
that host malware or viruses \cite{Canali2011,Mamun2016}.
There are also applications to online advertising
including anticipation of conversions \cite{Qiu2020}, contextual analysis of
the content\cite{Kan2005,Baykan2009,Meshkizadeh2010} or language\cite{Baykan2013} of webpages.

@@

A large amount of recent work on machine learning based on URLs has focused on deep learning approaches that
learn an embedding representation of the content of the URL. Once of the explicit reasons 
for this approach is the potential for sensitivity to the sequential structure of the URL.
\cite{Le2018}

Deep learning approaches include the use of convolutional layers to learn word and character 
embedings \cite{Le2018},

\cite{Ayes2019}
\cite{Li2020}
\cite{Basnet2012}
\cite{Basnet2014}
\cite{Tupsamudre2019}
\cite{Baykan2009}
\cite{Kan2005}
\cite{Hernandez2012}
\cite{Chung2010}
\cite{Qiu2020}
\cite{Xu2021}
\cite{Rhea2022}
\cite{Ma2009}
\cite{Verma2017}
\cite{Sountharrajan2020}
\cite{Murthy2022}

Classify Web pages based on URLs from the 4 Universities WebKB dataset. Comparing features derived from segmentation
of the URL to features derived from either link anchor text or the web page document itself \cite{Kan2004}.

Common approaches to feature engineering on URLs include string patterns and regular expressions to
identify key sequences and lexical properties \cite{Kan2004,Mamun2016}, 
creation of lookup tables or likelihood scores based on key sequences
\cite{Meshkizadeh2010}, n-gram or bag of words models \cite{Baykan2009}, 
the generation of task specific embedding vectors\cite{Le2018,Qiu2020} and the
usage of domain name servers or registrars for ancillary information about the domain registration \cite{Canali2011}. 
Some authors have 
emphasized that URLs are sequences of characters and require feature extraction methods that respect 
this sequential nature\cite{Le2018,Vazhayil2018}. They have naturally turned to developing neural network approaches
based on convolutional or recurrent layers for learning these sequential structures.
However, the ordering of key features within URLs is predominantly fixed as shown in Figure \ref{fig:url_structure}.
The core elements of URL structure occur at fixed positions within the sequence, in contrast to human language
sequences which are built from flexible grammars that allow variable positioning of most key elements.

\begin{figure*}
\centering
\includegraphics[scale=0.3]{images/URL_parts.png}
\caption{Structural components of a URL requiring specific feature engineering treatment}
\label{fig:url_structure}
\end{figure*}

@@

The features generated by the url2feature package are summairized in Table \ref{tab:features}

\begin{table*}
\caption{URL Features}
\label{tab:features}
\begin{tabular}{|l|l|l|l|}
\toprule
Group         &Feature              &Type        &Definition  \\
\midrule
Protocol      &protocol\_name         &Category    &The exact internet protocol name    \\
              &protocol\_type         &Category    &Internet protocol categorised by its purpose    \\
              &protocol\_exists       &Boolean     &Flag indicating presence or absence of internet protocol   \\
\midrule
Host          &host\_is\_ip           &Boolean     &Flag indicating if the host is an IP address    \\
              &host\_has\_port        &Boolean     &Flag indicating if a port number is explicitly specified    \\
              &domain\_sections       &Numeric     &Count of period separated domain sections    \\
              &domain\_reg\_year      &Numeric     &Year in which the domain was initially registered    \\
              &subdomain\_type        &Category    &Categorisation of the purpose of the subdomain \\
              &subdomain\_freq        &Numeric     &Frequency of the subdomain in internet traffics    \\
              &tld\_type              &Category    &Top Level Domain extension categorised by purpose    \\
              &tld\_freq              &Boolean     &Top Level Domain extension frequency in internet traffic   \\
\midrule
Path          &path\_depth            &Numeric     &The depth of the directory path between host and file   \\
              &path\_words            &Numeric     &Count of distinct words in the path between host and file   \\
              &path\_has\_date         &Boolean     &Flag indicating if a date is detected in the path  \\
              &path\_is\_home          &Boolean     &Flag indicating if the path starts with a home directory    \\
\midrule 
File          &file\_ext              &Category    &The exact file extension e.g. "exe" or "php"    \\
              &file\_type             &Category    &File extension categorised by function \& purpose    \\
              &file\_ext\_exists      &Boolean     &Flag indicating presence of a file extension for the URL target  \\
\midrule
Params        &params\_len            &Numeric     &The character length of the parameter string    \\
              &params\_count          &Numeric     &The count of the key value pairs in the param string    \\
              &params\_has\_url       &Boolean     &Flag indicating presence of a raw URL within the parameter string   \\
              &params\_enc\_url       &Boolean     &Flag indicating presence of an encoded URL within the parameter string   \\
              &params\_enc\_char      &Boolean     &Flag indicating presence of an encoded character within the parameter string   \\
              &params\_frag\_len      &Numeric     &Length of a hash delineated fragement at the end of the parameter string   \\
\bottomrule
\end{tabular}
\end{table*}

@@

The data used in this study was collected from a collction of publications that 
present internet classification problems containing URLs as the primary feature.
The ISCX Malcious URL classification dataset contains both malware and phishing URLS 
that need to be discriminated against benign URLS\cite{Mamun2016}.
The world wide web knowledge base (WebKb) 4 Universities data set contains a wide
range of university URLs categorised into multiple topic categories\cite{Craven1998}.
These 3 URL classification problems are summarised in Table \ref{tab:data}

The Syskill \& Webert webpage ratings dataset containing webepages across 4 categories
with \cite{Pazzani1996}

\begin{table}
\caption{Datasets}
\label{tab:data}
\begin{tabular}{|l|r|r|r|}
\toprule
Dataset              &Records        &Classes  &Majority    \\
\midrule
Malware              &46,944          &2       &            \\
Phishing             &45,343          &2       &            \\
WebKb 4Uni           &8,284           &7       &            \\
DMOZ                 &8,284           &7       &            \\
Syskill \& Webert     &8,284           &7       &            \\
\bottomrule
\end{tabular}
\end{table}

We generate all url features for each of these datasets and then apply a range of machine learning
techniques to build an appropriate classifier.
